Density dependence

### 2020-10-21 analysis/paper/appendix_density_dependence.Rmd
updated to use scaled-t family distribution

```{r data}
ha <- read_rds(here("analysis", "data", "derived_data", "ha_density.rds"))
# ha
```

@@

```{r}
g_growth <-
  gam(shts_next ~ 
        # s(ha_id_number, bs = "re") + #this doesn't seem to converge
        s(plot, bs = "re") +
        s(log_shts, bs = "cs") + #when modeled as a spline, edf is essentially a line
        # s(log_shts, bs = "cs", by = habitat) + #kind of like an interaction.
        habitat +
        s(Q, L,
          bs = "cb",
          # k = c(3,3),
          xt = list(bs = "cr")),
      family = gaussian(link = "log"),
      data = ha,
      method = "REML")
g_growth$smooth
```

@@

Where $f \cdot w(x, d)$ is a two dimensional crossbasis smooth composed of two marginal functions: $f(x)$, the effect of conspecific density, $x$ , on growth, and the additional distance response function $w(d)$ that models the effect of density at different annuli, $d$, equal to the midpoint of the inner and outer radius of the 5 annuli for which density was calculated. Each dimension was modeled with 10 knots. With a penalized spline, the number of knots is not very important as ... <!--# finish thought and cite (maybe Colin Edwards?) --> Habitat is a parametric fixed effect with two levels (1-ha and continuous forest), $s(log(shts_t))$ is a penalized cubic spline describing the growth from year t to year t + 1, and a random effect of plot, nested within habitat. We did not include a random effect of plant ID, as the model did not converge. Although the response, shoot number in year t+1, is a discrete variable, a log-normal distribution was used as it was a better fit than a poisson distribution ($\Delta$AIC = 527.2).

```{r}
g_growth_pois <- update(g_growth, family = poisson(link = "log"))
AICtab(g_growth, g_growth_pois)
```

@@

```{r}
g_surv <-
  gam(surv ~
        s(plot, bs = "re") +
        s(log_shts) +
        habitat +
        s(Q, L,
          bs = "cb",
          xt = list(bs = "cr")),
      family = binomial,
      data = ha,
      method = "REML")
```

@@

```{r}
# g_growth2 <- update(g_growth, ~-s(plot, bs = "re"))
# AICtab(g_growth, g_growth2)
# g_growth3 <- update(g_growth, ~-s(log_shts) + log_shts)
# AICtab(g_growth, g_growth2, g_growth3)
# Best model:
anova(g_growth)
```

>>>>

```{r data}
ha <- read_rds(here("analysis", "data", "derived_data", "ha_density.rds"))
ha$year %>% range()
```

@@

```{r}
g_growth <-
  gam(shts_next ~ 
        # s(ha_id_number, bs = "re") + #this doesn't seem to converge
        s(plot, bs = "re") +
        # s(log_shts, bs = "cr", by = habitat) + #prelim modeling shows this is a linear relationship (edf ~ 1)
        habitat * log_shts +
        s(Q, L,
          bs = "cb",
          # k = c(3,3),
          xt = list(bs = c("cr"))),
      family = gaussian(link = "log"),
      data = ha,
      method = "REML")
```

@@

Where $f \cdot w(x, d)$ is a two dimensional crossbasis smooth composed of two marginal functions: $f(x)$, the effect of conspecific density, $x$ , on growth, and the additional distance response function $w(d)$ that models the effect of density at different annuli, $d$, equal to the midpoint of the inner and outer radius of the 5 annuli for which density was calculated. Each dimension was modeled with 10 knots. With a penalized spline, the number of knots is not very important as ... <!--# finish thought and cite (maybe Colin Edwards?) --> Habitat is a parametric fixed effect with two levels (1-ha and continuous forest), $s(log(shts_t))$ is a penalized cubic spline describing the growth from year t to year t + 1, and a random effect of plot, nested within habitat. We did not include a random effect of plant ID, as the model did not converge. Although the response, shoot number in year t+1, is a discrete variable, a scaled t distribution with a log-link was used as it was a better fit than a poisson distribution ($\Delta$AIC = 810.5) and a log-normal distribution ($\Delta$AIC = 294).

```{r}
g_growth_pois <- update(g_growth, family = poisson(link = "log"))
g_growth_scat <- update(g_growth, family = scat(link = "log")) #because residuals are leptokurtic with gaussian
AICtab(g_growth, g_growth_pois, g_growth_scat)
```

@@

```{r}
g_surv <-
  gam(surv ~
        s(plot, bs = "re") +
        s(log_shts, bs = "cr", by = habitat) +
        habitat +
        s(Q, L,
          bs = "cb",
          xt = list(bs = "cr")),
      family = binomial,
      data = ha,
      method = "REML")
# plot(g_surv)
```

@@

```{r}
# g_growth2 <- update(g_growth, ~-s(plot, bs = "re"))
# AICtab(g_growth, g_growth2)
# g_growth3 <- update(g_growth, ~-s(log_shts) + log_shts)
# AICtab(g_growth, g_growth2, g_growth3)
# Best model:
anova(g_growth_scat)
# plot(g_growth_scat)
```
